---
title: "SimpleEDA Vignette"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{simpleEDA-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
linkcolor: blue
colorlinks: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette introduces the SimpleEDA R package and shows how to use it to perform
an Exploratory Data Analysis (EDA) of a data set. This package is intended to simplify
the EDA that a statistician would perform when analyzing a data set with the intention
of constructing a regression model. Please note that this package is called SimpleEDA 
not AutomaticEDA! The package presents the results, but does no interpretation, and 
may not perform all of the necessary tests based on the properties of the data set. 
For that, a trained statistician would be required.

Don't be scared by the size of this document. There is actually very little writing
or code. Most of the body is taken up by the detailed outputs and graphics from calls 
to the `simpleEDA` function.

While this vignette will show how to use the SimpleEDA package, it will not go into
the details of how to analyze the outputs or discuss the assumptions and results 
of the tests that are performed. For such details we recommend consulting a textbook
that covers linear regression models.

# Usage

Once the SimpleEDA package has been installed (not shown here), we load the SimpleEDA 
library:
```{r setup}
library(simpleEDA)
```

Next, load the desired data set as an R data frame. Here we use the `mtcars` 
data set that comes built in with R, and we set both the `am` and `vs` columns to
be factor variables.
```{r}
data(mtcars)

mtcars$am <- factor(mtcars$am, labels = c('automatic', 'manual'))
mtcars$vs <- factor(mtcars$vs, labels = c('v-shaped', 'straight'))
head(mtcars)
```

Once the data frame has been loaded into R, we can call the `simpleEDA` function
by giving it the data frame to analyze, the name of the response variable column,
and a list of the names of all the predictor variable columns.

The `simpleEDA` function can handle multiple predictor variables, and intelligently
chooses which tests to perform and which visualizations to generate based on whether
variables are continuous or categorical. The response variable can also be either
continuous or categorical.

## Single Continuous Predictor Variable

Suppose we wanted miles per gallon (`mpg`) to be the response variable, and horsepower (`hp`)
to be the only predictor variable. Then we would call `simpleEDA` like so: ([Jump to Summary](#cont_notes))

```{r}
simpleEDA(data_frame = mtcars, response_column_name = 'mpg',
          predictor_column_names = c('hp'))
```

### Summary {#cont_notes}
The `simpleEDA` function started by performing a univariate analysis for each of the two
continuous variables by printing distribution statistics, running the Shapiro-Wilk
normality test, and presenting histograms and boxplots of the distributions for
both variables along with their log-transforms. Next, it performed a correlation analysis
by showing a scatter plot of `hp` vs `mpg` with a Lowess smooth, computed the correlation,
and applied a Box-Cox transform of the model with `mpg` as the response and `hp` as the predictor.

## Single Categorical Predictor Variable

Next, let's see how the analysis changes if there is a single categorical predictor variable.
([Jump to Summary](#cat_notes))

```{r}
simpleEDA(data_frame = mtcars, response_column_name = 'mpg',
          predictor_column_names = c('am'))
```

### Summary {#cat_notes}
The univariate analysis for `mpg` is the same as before, but looking at the analysis
for the `am` column, we can see that it now prints a frequency table and bar chart
showing the number of occurrences in each category of the `am` variable. 

Next, in the section analyzing the correlation between the continuous response and 
the categorical predictor, the `simpleEDA` function presents box plots of the distribution 
of mpg within each category, a conditional density plot, the results of the Shapiro-Wilk
normality test applied to the distribution of `mpg` results in each category of `am`,
Cohen's d test, an ANOVA t-test, and the Kruskal-Wallis test.

## Full Model

Lastly, we can examine what happens when applying `simpleEDA` with multiple predictor
variables, some categorical and some continuous. Let's say that we wanted miles 
per gallon (`mpg`) to be the response variable, and horsepower (`hp`), 
weight (`wt`), cylinder shape (`vs`), and automatic/manual (`am`) to be the predictor variables.
We would call `simpleEDA` like so: ([Jump to Summary](#full_notes))

```{r}
simpleEDA(data_frame = mtcars, response_column_name = 'mpg',
          predictor_column_names = c('hp', 'wt', 'vs', 'am'))
```

### Summary {#full_notes}
The same kinds of univariate analyses are performed, and the comparisons
between the response variable and the predictors show the same types of output as in
the previous sections.

In addition to those, `simpleEDA` now presents results that compare each of the predictor
variables to one another so that they can be analyzed for multicollinearity. Note
that in all of these comparisons, the visualizations and tests are intelligently chosen
based on whether the variables are continuous or categorical.

